---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(tidyverse)
library(dplyr)
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 
```{r}
deaths <- av %>%
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
               filter(Death != "")

deaths$Time <- parse_number(deaths$Time)

head(deaths)
```

Similarly, deal with the returns of characters.
```{r}
returns <- deaths %>%
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time_returns",
               values_to = "Return") %>%
               filter(Return != "")

returns$Time_returns <- parse_number(returns$Time_returns)

head(returns)
```

Based on these datasets calculate the average number of deaths an Avenger suffers.
```{r}
avg <- colSums(returns == "YES")
avg[13]/173
```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

"There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife"

### Include the code

```{r}
firsts <- subset(returns, returns$Time == 1)
firsts <- subset(firsts, firsts$Time_returns == 1)
firstreturn <- subset(firsts, firsts$Return == "YES")
count(firstreturn) / count(firsts)

```


### Include your answer

When it comes to return rate of first time deaths for the Avengers, there is exactly a 2 in 3 chance that if you die, you will return. This was found by narrowing the results to first time deaths and returns and then dividing the count of people who returned by the total count of first deaths. 

Upload your changes to the repository. Discuss and refine answers as a team.

### FiveThirtyEight Statement

**The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times.**

### Include the code

```{r}
filter(returns, Name.Alias == "Jocasta")[13:15]
```

### Include your answer

**This data shows that Jocasta has indeed died 5 times and returned each time, which is consistant with the quote from the FiveThirtyEight article. This was done by filtering out the rows with "Jocasta" as the Name.Alias, and then finding the columns with Death count and Time in them.**

